### 2015-08-26 doc/manual/runs.tex

The magnetic drift terms introduce nonzeros in the system matrix,
and therefore increase the memory required for factorization; a typical increase is 20-40\%.
These magnetic drift terms typically have a minor effect except when the radial
electric field is near 0.

@@

\section{Parallelization}
[Coming soon...] Choosing the number of nodes and procs.

\section{Issues with running on 1 processor}

>>>>

The magnetic drift terms introduce nonzeros in the system matrix,
and therefore increase the memory and time required for factorization.
The change is small; a typical increase in both memory and time is 20-40\%.
These magnetic drift terms typically have a minor effect on the physics outputs except when the radial
electric field is near 0.

@@

\section{Sparse direct solver packages}
\label{sec:solvers}

As discussed briefly in section \ref{sec:gmres}, the most computationally demanding step in 
\sfincs~is the direct $LU$-factorization of a very large sparse nonsymmetric real matrix. The \PETSc~library
which \sfincs~uses has interfaces to a large number of other packages for direct factorization of such matrices,
making it possible to choose among the various solver packages with just a command-line flag ({\ttfamily -pc\_factor\_mat\_solver\_package}).  Some lists of the direct solvers
available in \PETSc~can be found \href{http://www.mcs.anl.gov/petsc/petsc-current/docs/manualpages/Mat/MatSolverPackage.html#MatSolverPackage}{here}
or in the ``direct solvers''-``LU'' section of \href{http://www.mcs.anl.gov/petsc/documentation/linearsolvertable.html}{this page}.
It is important for the $LU$-solver package to be one that is efficiently parallelized, 
in order to be able to solve problems at the high resolutions required for experimentally relevant collisionality and magnetic geometry.
The recommended choice of $LU$ solver is \href{http://mumps-solver.org/}{\mumps} (which is the default), and another good option is \href{http://crd-legacy.lbl.gov/~xiaoye/SuperLU/}{\superludist}.
(The {\ttfamily PARDISO} library available in the Intel Math Kernel Library is probably suitable as well, though we have not investigated it yet.)
In side-to-side comparisons, we find \mumps~systematically uses substantially less memory and time than \superludist~for factorization.
In principle, other solver packages for asymmetric matrices that are interfaced to \PETSc~could be used as well, such as {\ttfamily UMFPACK}, {\ttfamily PASTIX}, etc.

There are two ways to choose between solver packages.
One method is the \sfincs~parameter \parlink{whichParallelSolverToFactorPreconditioner} in the 
{\ttfamily \hyperref[sec:otherNumericalParameters]{otherNumericalParameters}} namelist.
This parameter only allows you to choose between \mumps~ and \superludist, not other solver packages interfaced to \PETSc.
Another way to choose between solver packages is the command-line flag {\ttfamily -pc\_factor\_mat\_solver\_package},
followed by one of the options in quotation marks \href{http://www.mcs.anl.gov/petsc/petsc-current/docs/manualpages/Mat/MatSolverPackage.html#MatSolverPackage}{here}.
The command-line flag overrides the namelist parameter.

The physics outputs of the code should be independent of the solver package used to several significant digits.
In principle, different solver packages solving the same linear system should find the identical solution.
However there will be small differences in the solutions associated with roundoff error.

Note that {\ttfamily superlu} and \superludist~are distinct libraries. The former is serial while the latter is parallelized.
Therefore there is no reason to use {\ttfamily superlu}; \superludist~is always preferable.

The \mumps~package has a large number of control parameters, which are documented in the \mumps~manual which can be downloaded \href{http://mumps-solver.org/}{here}.
You do not need to be aware of most of these control parameters.
However, several parameters which may be useful are discussed in section \ref{sec:mumpsControlParameters}.
It is also worth being aware of the section of the \mumps~manual on error messages. 
This section is useful for interpreting the {\ttfamily INFO(1)} and {\ttfamily INFO(2)} error codes that
are reported if \sfincs~exits with an error associated with \mumps.

The \superludist~package has many fewer options than \mumps. 
You can find of list of the options by running \sfincs~with the {\ttfamily -help} command-line flag when
using \superludist, and searching the output for lines containing {\ttfamily superlu\_dist}.
The \superludist~options are also documented in the package's manual, available \href{http://crd-legacy.lbl.gov/~xiaoye/SuperLU/#superlu_dist}{here}.
We have not found any advantage in adjusting
any of the \superludist~options.

The \PETSc~library includes a built-in sparse direct solver which works on only a single processor.
You can select this solver using the command-line flag \\
\centerline{\ttfamily -pc\_factor\_mat\_solver\_package petsc}\\
This solver could potentially be useful if you are running on a system that does not have \mumps~or \superludist~installed,
and you are only considering problems that require sufficiently little memory (e.g. tokamaks) that parallelization is not required.
However, this solver is less robust than \mumps~or \superludist, sometimes exiting with an error message that there is a zero pivot
even though  \mumps~and \superludist~can solve the same system with no problem. Therefore, even if you plan to use only a single processor,
we still recommend that you install  \mumps~or \superludist.

\section{Parallelization: Choosing the number of nodes \& processors}
\label{sec:parallelization}

Usually the limiting factor for \sfincs~is not time but memory.
(The time required depends on the resolution used, but jobs for experimentally relevant W7-X parameters
typically take under 10 minutes.)
Therefore, when considering how many nodes to request for a \sfincs~job,
the first issue to consider is ensuring you have requested sufficient total memory.
A good way to determine the memory required (assuming you are using the default solver \mumps)
is to first run \sfincs~on 1 node using the parameters of interest
and look for the following line in standard output:\\
\centerline{\ttfamily ** TOTAL     space in MBYTES for IC factorization         :       1072}\\
The number at the end will generally be different; it depends not only on the
resolution and number of species used, but also increases with the number of processors
as discussed below.  Make sure the number of nodes requested times the number of megabytes per node exceeds this number.
This line in standard output is generated by \mumps, so if you are using a different solver, this information is not printed,
and you may need to determine the number of nodes by trial-and-error.
Since some memory on each node is used by the operating system and by \sfincs~functions other than the solver,
you may need to use a slightly higher number of nodes than this estimate suggests.
For experimentally relevant W7-X and HSX parameters, we typically use 2-6 nodes with 64 GB each.
Problems with lower resolution requirements, such as tokamaks, often can be run on a single node.

The `IC' in the above line stands for `in core', meaning the $L$ and $U$ factors are stored
in memory rather than on disk.  In \mumps, one can also choose to do an `out of core' (OOC) solve,
in which case substantially less memory is typically required.  The price you pay for this memory savings is time, since
disk access is slow compared to memory access.  Due to this slowdown, we have not used the OOC capability much, but
you might find it useful in some circumstances.
To see how much memory would be required for an OOC solve,
look for the following line in standard output:\\
\centerline{\ttfamily ** TOTAL     space in MBYTES for OOC factorization        :       128}\\
(The number at the end will generally be different.)
To invoke out-of-core mode, you must take two steps. First, use the following command-line flag when calling \sfincs:\\
\centerline{\ttfamily -mat\_mumps\_icntl\_22 1}\\
Second, you must set the environment variable {\ttfamily MUMPS\_OOC\_TMPDIR} to some reasonable directory
before calling \sfincs. This environment variable could be set for example in the batch job file.
Temporary files containing the $L$ and $U$ factors will be stored in the directory indicated.

Another issue to consider is how many processors to use.  It is not always best to use the maximum number of processes
available on the number of nodes you have chosen. The reason is that as the preconditioner matrix is divided among
more and more processes, the $LU$ factorization becomes less efficient, requiring more memory and more communication.
If you examine the \mumps~IC and OOC memory requirements indicated above, you will find they increase somewhat 
as the number of processes increase.
One needs to find a balance between speed (favoring many processes) and memory requirements (favoring few processes).
While it is almost always better overall to use 2 processes compared to 1 process, it is not always better to use 128 processes
compared to 64 processes.  The sweet spot is often in the range of 16-64 processes.
To determine how to request fewer processes than the maximum available on a given number of nodes,
see the documentation for your computing system.

If more memory is required than is available, the system will usually terminate your job with an out-of-memory (OOM) error.
When this occurs, you need to either increase the number of nodes requested or decrease the number of processors
requested.

Most of the time, \sfincs~is run via \sfincsScan~as some parameter is scanned, such as the radial electric field.
In this case, the scan is ``embarassingly parallel'' in the sense that each job in the scan is completely independent
of the other jobs.  Even if each individual job requires only 1 or a small number of nodes, it is still useful
to run \sfincs~on a computing system with many nodes so the scan can be carried out in parallel.

%\section{Issues with running on 1 processor}
